---
title: "Decision Making Under Uncertainty, MATH 20011"
subtitle: "Homework 2: Hypothesis Testing"
author: "Christian Auman"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

* Replace "Your Name Here" in the author field above with your name.

* Follow the prompts in each part of this assignment and edit this .Rmd file directly. There are four parts to this assignment.

* Knit your finished document as PDF and submit the PDF on Canvas.


# Libraries

```{r library, echo = TRUE, message = FALSE, warning = FALSE}
library(openintro) # Library required to access our dataset
library(ggplot2) # Library required for visualization
library(dplyr) # Library required for data tidying and manipulation
```


# PART 1: Data Exploration 

We will use the `heart_transplant` dataset from the `openintro` package for this lab.

(a) Load the specified data set. Hint: use `data()`

```{r}
data(heart_transplant)
```

(b) Compactly display the internal structure of the data. Hint: use `str()`

```{r}
str(heart_transplant)
```

(c) View the first 6 rows of your data. Hint: use `head()`

```{r}
head(heart_transplant, 6)
```

(d) Replace `eval=FALSE` with `eval=TRUE`. Explain the purpose and output of the following R code. 

- **Eval determines whether the code is executed in the PDF render**

```{r eval=TRUE}
heart_transplant |> filter(!is.na(wait))
```

**Question:**

Based on your observations from the (a) - (d) and the information provided by (openintro)[https://www.openintro.org/data/index.php], describe the `heart_transplant` dataset. Your description must include, but not limited to:

- How many observations are in this dataset?

- What variables are included? Identify which are categorical and which are numerical.

- Why some entries in the `wait` column are `NA`?

**Answer:**

1. There are 103 observations in the dataset
2. There are 8 variables:
    - id - numeric continuous
    - acceptyear - numeric discrete 
    - age - numeric continuous
    - survived - categorical nominal
    - survtime - numerical discrete
    - prior - categorical nominal
    - transplant - categorical nominal
    - wait - numerical discrete
3. Some entries in the wait columnn are N/A because some few candidates were showing signs of improvement and they got deselected as a candidate, but they were kept in the data. And because they never got a transplant, there wait time becomes N/A

\newpage

# Part 2: Data Summary and Visualization

(a) Generate a summary statistics of the data. Hint: use `summary()`

```{r}
summary(heart_transplant)
```

(b) Create a bar chart showing the number of patients by treatment. Hint: use the `transplant` variable. 

```{r}
vals <- table(heart_transplant$transplant)
barplot(height = vals, xlab = "Transplant", ylab = "Patients", main = "Number of patients by treatment", names.args = names(vals))
```

(c) Create a histogram of patient ages. Hint: use the `age` variable. 

```{r}
hist(heart_transplant$age, xlab = "Age")
```

(w) Create a boxplot comparing survival time (`survtime`) by treatment (`transplant`) group.

```{r}
boxplot(
    heart_transplant$survtime[heart_transplant$transplant == "control"],
    heart_transplant$survtime[heart_transplant$transplant == "treatment"],
    names = c("Control", "Treatment")
)
```

(e) Create a scatterplot of `age` vs. `survtime` with color as `transplant`.

```{r}
colors <- ifelse(heart_transplant$transplant == "treatment", "blue", "red")
plot(
    heart_transplant$age,
    heart_transplant$survtime,
    col = colors
)
```

**Questions:**

Based on your observation from (a) - (e), describe your data using the following prompts: 

- The distribution of the age.

- The group (treatment/control) with the longest median survival time.

- Relationship between age and survival time.

**Answer:**

1. From the histogram, the majority of the patients fall between the ages of 40-60
2. Given the box plot, the group with the longest median survival time were the treatment group
3. In the scatterplot, the patients between the ages of 30-50 had longer survival times. But I would have to say that I cannot generalize age and survival time, as the sample size is small, and the groups were mainly between the ages of 40-60 which is not representative of the population.


\newpage

# Part 3: Hypothesis Test for a Proportion

(a) Create a binary variable indicating whether a patient survived more than 30 days post-transplant. Store the new tibble in a new tibble `heart_transplant2`

```{r}
heart_transplant2 <- heart_transplant |> mutate(survtime_30 = ifelse(heart_transplant$survtime > 30, 1, 0))
heart_transplant2
```

(b) Replace `eval=FALSE` with `eval=TRUE`. Performing hypothesis test for one population proportion. 

```{r eval=TRUE}
x <- sum(heart_transplant2$survtime_30)
n <- nrow(heart_transplant2)
prop.test(x = x, n = n, p = 0.5, alternative = "two.sided", conf.level = 0.95)
```

**Questions:**

- Explain each line of code within the code chunk in (b)

**Answer:**

`x <- sum(heart_transplant2$survtime_30)`
- This gets the number of patients who survived more than 30 days 
`n <- nrow(heart_transplant2)`
- This is the number of total patients

> They are both used to calculate the sample proportion, x/n

`prop.test(x = x, n = n, p = 0.5, alternative = "two.sided", conf.level = 0.95)`
- This is testing the null hypothesis and an alternative hypothesis. The null hypothesis is whether true proportion is equal to 0.5, and tests the alternative hypothesis where proportion is not equal to 0.5
```{r echo=FALSE}
prop.test(x = x, n = n, p = 0.5, alternative = "two.sided", conf.level = 0.95)
```

Here the p-value is much less than 0.05 meaning that the hypothesis where proportion is 0.5 is rejected, having the true proportion != to 0.5. The sample proportion is around 0.767 with a 95% CI

- What is the sample proportion of patients who survived beyond 30 days?

**Answer:**

The sample proportion is `0.7669903`:
```{r}
x / n
```

- Interpret the $95\%$ confidence interval in context of the study.

**Answer:**

We are 95% confident that the true proportion lies between `0.67` and `0.84`

- State the null and alternative hypotheses for this test.

**Answer:**

The null hypothesis is when proportion is = 0.5
The alternative hypothesis is when proportion is != 0.5

The proportion here is the number of patients who survived longer than 30 days vs the total sample size

- What is the p-value? 

**Answer:**

Assuming that the null hypothesis is true, the p-value represents the probability of producing a sample proportion which is as far from 0.5 as our sample proportion

- Do you reject or fail to reject the null hypothesis at $\alpha = 0.05$ and why?

**Answer:**

I Reject the null hypothesis as the p-value is much less than 0.05, which means the true sample proportion is more likely not equal to 0.5

- Interprete the result of the hypothesis test.

**Answer:**

I would conclude given the hypothesis test, that the proportion of patients who survived longer than 30 days is more likely not 50%


\newpage 

# Part 4: Hypothesis Test for a Mean

In this part, we are interested in knowing whether the average age of heart transplant patients is different from 50 years?

Replace `eval=FALSE` with `eval=TRUE` to see the result from the code chunk.

```{r eval=TRUE}
t.test(heart_transplant2$age, mu = 50, alternative = "two.sided", conf.level = 0.95)
```


**Questions:** Provide answers to the following questions:

- Explain the line of code within the code chunk in this part

**Answer:**

`t.test` is compares the mean instead of proportion. Here, i'ts comparing the age mean to 50. It's "two.sided" which tests whether the true mean is not 50

- What is the sample mean age of the patients?

**Answer:**

The sample mean age of the patients is 44.64078 years old

- Interpret the $95\%$ confidence interval in context of the study.

**Answer:**

We are 95% confident that the true age mean of the population lies between `42.73` and `46.56`

- State the null and alternative hypotheses for this test.

**Answer:**

The null hypothesis was if mean age was equal to 50
The alternative hypothesis was if mean age was not equal to 50

- What is the p-value? 

**Answer:**

Assuming the null hypothesis is true, the p-value represents the probability of producing a mean age as far from 50 as our sample age mean

- Do you reject or fail to reject the null hypothesis at $\alpha = 0.05$ and why?

**Answer:**

I reject the null hypothesis because the p-value was much less than 0.05, meaning the true mean is more likely not 50

- Interprete the result of the hypothesis test.

**Answer:**

The hypothesis test shows that the true mean age of the populationo is more likely not 50. We are instead 95% confident that the true mean lies between 42 and 46.




