---
title: "Decision Making Under Uncertainty, MATH 20011"
subtitle: "Homework 2: Hypothesis Testing"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

* Replace "Your Name Here" in the author field above with your name.

* Follow the prompts in each part of this assignment and edit this .Rmd file directly. There are four parts to this assignment.

* Knit your finished document as PDF and submit the PDF on Canvas.


# Libraries

```{r library, echo = TRUE, message = FALSE, warning = FALSE}
library(openintro) # Library required to access our dataset
library(ggplot2) # Library required for visualization
library(dplyr) # Library required for data tidying and manipulation
```


# PART 1: Data Exploration 

We will use the `heart_transplant` dataset from the `openintro` package for this lab.

(a) Load the specified data set. Hint: use `data()`

```{r}

```

(b) Compactly display the internal structure of the data. Hint: use `str()`

```{r}

```

(c) View the first 6 rows of your data. Hint: use `head()`

```{r}

```

(d) Replace `eval=FALSE` with `eval=TRUE`. Explain the purpose and output of the following R code. 

```{r eval=FALSE}
heart_transplant |> filter(!is.na(wait))
```

**Question:**

Based on your observations from the (a) - (d) and the information provided by (openintro)[https://www.openintro.org/data/index.php], describe the `heart_transplant` dataset. Your description must include, but not limited to:

- How many observations are in this dataset?

- What variables are included? Identify which are categorical and which are numerical.

- Why some entries in the `wait` column are `NA`?

**Answer:**


\newpage

# Part 2: Data Summary and Visualization

(a) Generate a summary statistics of the data. Hint: use `summary()`

```{r}

```

(b) Create a bar chart showing the number of patients by treatment. Hint: use the `transplant` variable. 

```{r}

```

(c) Create a histogram of patient ages. Hint: use the `age` variable. 

```{r}

```

(d) Create a boxplot comparing survival time (`survtime`) by treatment (`transplant`) group.

```{r}

```

(e) Create a scatterplot of `age` vs. `survtime` with color as `transplant`.

```{r}

```

**Questions:**

Based on your observation from (a) - (e), describe your data using the following prompts: 

- The distribution of the age.

- The group (treatment/control) with the longest median survival time.

- Relationship between age and survival time.

**Answer:**



\newpage

# Part 3: Hypothesis Test for a Proportion

(a) Create a binary variable indicating whether a patient survived more than 30 days post-transplant. Store the new tibble in a new tibble `heart_transplant2`

```{r}

```

(b) Replace `eval=FALSE` with `eval=TRUE`. Performing hypothesis test for one population proportion. 

```{r eval=FALSE}
x <- sum(heart_transplant2$survtime_30)
n <- nrow(heart_transplant2)

prop.test(x = x, n = n, p = 0.5, alternative = "two.sided", conf.level = 0.95)
```

**Questions:**

- Explain each line of code within the code chunk in (b)

**Answer:**

- What is the sample proportion of patients who survived beyond 30 days?

**Answer:**

- Interpret the $95\%$ confidence interval in context of the study.

**Answer:**

- State the null and alternative hypotheses for this test.

**Answer:**

- What is the p-value? 

**Answer:**

- Do you reject or fail to reject the null hypothesis at $\alpha = 0.05$ and why?

**Answer:**

- Interprete the result of the hypothesis test.

**Answer:**



\newpage 

# Part 4: Hypothesis Test for a Mean

In this part, we are interested in knowing whether the average age of heart transplant patients is different from 50 years?

Replace `eval=FALSE` with `eval=TRUE` to see the result from the code chunk.

```{r eval=FALSE}
t.test(heart_transplant2$age, mu = 50, alternative = "two.sided", conf.level = 0.95)
```

**Questions:** Provide answers to the following questions:

- Explain the line of code within the code chunk in this part

**Answer:**

- What is the sample mean age of the patients?

**Answer:**

- Interpret the $95\%$ confidence interval in context of the study.

**Answer:**

- State the null and alternative hypotheses for this test.

**Answer:**

- What is the p-value? 

**Answer:**

- Do you reject or fail to reject the null hypothesis at $\alpha = 0.05$ and why?

**Answer:**

- Interprete the result of the hypothesis test.

**Answer:**








